---
title: "Environmental readings and detections"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Environmental readings and detections}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{css, echo=FALSE}
pre {
  overflow-x: scroll !important;
}
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# Initial Setup and Api Credentials
Please see [station_and_receptors](../articles/station_and_receptors.html) for further details about how to make available the showcased package `temploapiclient` to this vignette.
```{r setup}
project_folder <- dirname(getwd())
path_to_env <- file.path(project_folder, ".env")
readRenviron(path_to_env)
api_base_url <- Sys.getenv("API_BASE_URL")
```

```{r instantiate_client}
templo_client <- temploapiclient::TemploApiClient$new(api_base_url)
library("magrittr") # Make '%>%' available throughout the vignette
```
# Fetching environmental readings
Deployed receptors are equipped with sensors that measure oceanographic variables such as temperature, pressure, etc... Such readings can be obtained from the API too. Readings are generate with a very fine time resolution, so it is advised to fetch two o three days worth of data at the time. Otherwise, the API call may time out and result in an error.

In the example below we are fetching three days worth of data for 10th, 11th and 12th October 2024. We are just selecting and renaming a few fields for showcasing purposes.
```{r available_environmental_variables}
api_data <- templo_client$get_environmental_readings(from_date = "2024-10-10", to_date = "2024-10-12")
head(api_data) %>%
  dplyr::rename(c(
    variable = environmental_variable_name,
    receptor = receptor_name,
    station = station_name,
    lat = latitude,
    lon = longitude
  )) %>%
  dplyr::select(
    station,
    receptor,
    lat,
    lon,
    variable,
    reading_time,
    value
  )
```
We could hit an interval of time when no data was produced so an empty dataset will be returned instead.
```{r unavailable_environmental_variables}
api_data <- templo_client$get_environmental_readings(from_date = "2024-01-01", to_date = "2024-02-02")
api_data %>%
  assertr::verify(!nrow(.))
```
# Fetching detections
Detections emitted by the sensors fitted in the tags that are surgically inserted into individuals read oceanographic variables too. Similar to environmental readings, detections should be downloaded in small chunks to avoid the API timing out and producing no results.
In the results below we have massaged the data to provide contextual column names and trimmed those detections that produced no value.

```{r available_detections}
api_data <- templo_client$get_dataframe_end_point(end_point = "detections/data_exports/", parameters = list(
  format = "json", # this parameters, for our purposes needs no change.
  from_date = "2024-10-02",
  to_date = "2024-10-03"
))
api_data %>%
  dplyr::mutate(detection_value = as.numeric(detection_value)) %>%
  dplyr::filter(assertr::not_na(detection_value)) %>%
  dplyr::select(
    sensor_id,
    sensor_name,
    individual_tag,
    tag_model,
    scientific_name,
    station_name,
    receptor_name,
    detection_time,
    detection_value
  ) %>%
  dplyr::rename(c(
    id = sensor_id,
    sensor = sensor_name,
    individiual_id = individual_tag,
    station = station_name,
    receptor = receptor_name,
    time = detection_time,
    value = detection_value
  )) %>%
  head()
```
